---
title: "Week 8 A"
author: "Wali Reheman"
date: "2024-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Dataset**

We will use the **NHANES** dataset, which contains health and demographic information collected from a representative sample of the U.S. population. The dataset is available in R through the `NHANES` package.

**Accessing the Dataset**:

```{r}
install.packages("NHANES")  # Install if not already installed
library(NHANES)
data <- data.frame(NHANES)

```

## TASK 1：Data Exploration

-   **Load the `NHANES` dataset**.

-   Use `tidyverse` functions to:

    -   View the **structure** of the dataset.

    -   Summarize key variables: `Age`, `Gender`, `BMI`, `SmokeNow` (whether the person currently smokes), and `PhysActive`(physical activity status).

    -   Just check the **levels** if it is categorical variable.

```{r}


```

## TASK 2: Data Cleaning 

-   Check for **missing values** in key variables.

-   Remove any rows with missing values.

-   Recode `SmokeNow` and `PhysActive` to have more interpretable categories if necessary.

```{r}


```

## TASK 3: Data Visualization

-   Create a **histogram** of `BMI`.

-   Create a **boxplot** of `BMI` by `Gender`.

-   Add appropriate **labels** and **titles**.

```{r}

```

## TASK 4: **Descriptive Statistics**

-   Present sample statistics for our key variables, mutate the data if needed ( for categorical variables)

-   Present the results in a **neat table**.

```{r}

```

## TASK 5: Regression Analysis

-   Run a linear regression with `BMI` as the **dependent variable** and `Age`, `Gender`, and `PhysActive` as **independent variables**.

-   **Interpret** the coefficients ( to yourself or group member).

```{r}

```

## TASK 6: Presenting Results

-   Use `stargazer` to create a **regression table (both in the console and a Word doc)**.

-   Use `coefplot` to **visualize** the coefficients.

```{r}

```
