---
title: "Week 8 B"
author: "Wali Reheman"
date: "2024-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Dataset**

We will use the **NHANES** dataset, which contains health and demographic information collected from a representative sample of the U.S. population. The dataset is available in R through the `NHANES` package.

**Accessing the Dataset**:

```{r}
install.packages("NHANES")  # Install if not already installed
library(NHANES)
data <- data.frame(NHANES)

```

## **TASK 1: Data Exploration and Cleaning**

-   Load the `NHANES` dataset.

-   Check for **missing values** and handle them appropriately.

-   Summarize key variables: `Age`, `Gender`, `BMI`, `SmokeNow` (whether the person currently smokes), `PhysActive`(physical activity status) and  `SleepHrsNight`

-   Identify and remove **outliers** in `BMI` (e.g., using the IQR method).

-   **Transform** `BMI` using log transformation if you think it is appropriate.

-   Recode `SmokeNow` and `PhysActive` to have more interpretable categories if necessary.

```{r}


```

## TASK 2: Data Visualization

-   Create a **scatter plot** of `BMI` vs. `Age`.

-   Color the points by `SmokeNow`.

-   Use **smooth lines** to show trends (e.g., using `geom_smooth`).

-   Customize the plot with **themes** and **labels**.

```{r}

```

## TASK 3: **Descriptive Analysis**

-   Calculate **correlation coefficients** between `Age`, `BMI`, and `SleepHrsNight` (average hours of sleep per night, if available).

-   Present the **correlation matrix**.

```{r}
# Ensure 'SleepHrsNight' is available and has no missing values

```

## TASK 4: Multiple Regression Analysis

Run multiple regression models:

-   **Model 1**: `BMI ~ Age`

-   **Model 2**: Add `Gender` as an independent variable.

-   **Model 3**: Include `PhysActive` , `SmokeNow` and `SleepHrsNight` as independent variables.

```{r}


```

## TASK 5 (Optional) : Model Diagnostics

-   Check the **assumptions** of linear regression for `model3`:

    -   **Linearity**

    -   **Independence**

    -   **Homoscedasticity**

    -   **Normality of residuals**

```{r}



```

## TASK 6: Presenting Results

-   Use `stargazer` to **compare all four models** in a single table with **robust standard errors**.

-   Use `coefplot` to **visualize and compare** coefficients across models.

```{r}

```
